\addsec{Abstract}
\label{sec:zusammenfassung}

Machine Learning models, influencing almost every aspect of your daily lives without us noticing, have become a cornerstone of modern day amenities. When trusting these models to draw a decision based on causality instead of recalling from memory, statistical inference through stochastic processes provides the solution. Since reality is not as simple as our usual models assume, extracting the useful information from an array of noisy data becomes more important by the day. A financial market is therefore an optimal object of interest to study Bayesian models, which have proven useful in determining factors to describe such a system thoroughly. In this thesis a list of models able to infer important information from noisy systems is presented in order to find a solution to the observed systematical impairment observed in the previous models. They employ Gaussian and Student-t distributions to base the stochastic processes on, showing that the Gaussian and Student-t Process Latent Variable Model provide excellent reconstruction of the covariance structure of stock market data. The analysis of the systematical impairment indicate, that the covariance structure is above all a matter of the size of observed quantities. The work in this thesis therefore provides an arrangement of models with which market analysis can be improved upon through better understanding systematical influences of different variables on the results and interpretability of statistical latent variable models.